---
title: "RA_test_ARSHAD_AZAD"
author: "arshad"
email: "x@x.com"
date: "11/01/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Required packages}
library(tidyverse, warn.conflicts = F)
library(haven)
library(naniar)
library(ggplot2)
library(ggpubr)

```
# Part 1

```{r Part I}

```

## 1. Importing
```{r 1. Importing}
# a. Importing 'New Variables.csv' as new_variables

new_variables <- read.csv("~/Desktop/RA_test/coding_challenge/Part 1/New Variables.csv")


# b. Importing 'Main Dataset.dta' as main_dataset

main_dataset <- read_dta("~/Desktop/RA_test/coding_challenge/Part 1/Main Dataset.dta")

## merging new_variables with main_dataset

main_dataset <- main_dataset %>% left_join(new_variables, by = "uniqueid")

# c. Importing 'New Observations.dta' as new_observations

new_observations <- read_dta("~/Desktop/RA_test/coding_challenge/Part 1/New Observations.dta")

## adding new_observation to main_dataset

main_dataset <- bind_rows(main_dataset, new_observations)
```

## 2. Quality checks
### a. average and median values of time spent surveying for completed surveys

```{r 2. Quality checks}

main_dataset$surveytime <- as.numeric(main_dataset$surveytime)

avg_time <- mean(main_dataset$surveytime[main_dataset$survey_complete == 1])
paste0("average time spent is ", avg_time)


median_time <- median(main_dataset$surveytime[main_dataset$survey_complete == 1])
paste0("median time spent is ", median_time)
```
### b. survey time variations among surveyors

```{R}

df_survey_time <- main_dataset %>% select(surveyor, survey_complete, surveytime) %>% filter(survey_complete == 1) %>% group_by(surveyor) %>% mutate(avg_surveyor_time = mean(surveytime))

avg_time_for_surveyors <- unique(paste0("Surveyor ", df_survey_time$surveyor, " spent ", df_survey_time$avg_surveyor_time, " seconds."))

avg_time_for_surveyors
```

### c. Duplicate 'hhid'.


```{R}
## First we make a table with frequecies for each hhid
n_occur <- data.frame(table(main_dataset$hhid))

## Following gives the hhid with more than 1 entries
n_occur[n_occur$freq > 1,]

```
This subsets and then finds the 'uniqueid' for which 'hhid' are repeated, i.e, duplicate 'hhid' exists.

```{R}
main_dataset[main_dataset$hhid %in% n_occur$Var1[n_occur$Freq > 1],]$uniqueid
```

We can resolve the duplicate id problem as follows. Make a composite 'hhid2' from 'hhid' and 'uniqueid'. This gives a unique 'hhid2'.

```{R}

main_dataset$hhid2 <- paste0(main_dataset$uniqueid,main_dataset$hhid)

length(main_dataset$hhid2)
```

### d. Missing value plot
```{R}
# We select following 4 variables.

missing_data <- main_dataset %>% select(cycleownnum, cycleownvalue, own2wheelernum, own2wheelervalue) %>% miss_var_summary()

missing_data

```


```{R}

missing_data_plot <- ggbarplot(missing_data, x = "variable", y = "pct_miss", add = "none") + stat_summary(aes(label=round(..y..,2)), fun=mean, geom="text", size=3,
             vjust = -0.5) + ggtitle("Graph of missing data for some variables") + theme(plot.title = element_text(size = 10, face = "bold"))

missing_data_plot
ggsave("missing_data_plot.png")

```
### e. histograms and k-density plots


```{R}

ggplot(main_dataset, aes(surveytime)) +
  geom_histogram(aes(y=..density..)) +
  geom_density(col = "red") +
  ggtitle("Histogram for survey time", subtitle = "with an overlay of k-denstiy")

ggsave("hist_survey_time.png")

ggplot(main_dataset, aes(income)) +
  geom_histogram(aes(y=..density..)) +
  geom_density(col = "red") +
  ggtitle("Histogram for income", subtitle = "with an overlay of k-denstiy")

ggsave("hist_income.png")

```
## 3. Cleaning

### a. removing personally identifiable information
```{r 3. cleaning}



main_dataset$surveyor <- str_replace_all(main_dataset$surveyor, c("Benjamin" = "1", "Peter" = "2", "Anna" = "3", "Mary" = "4", "John" = "5", "Caroline" = "6", "Joseph" = "7", "Grace" = "8", "David" = "9", "Sam" = "10", "Jane" = "11"))

main_dataset$surveyor <- as.numeric(main_dataset$surveyor)

unique(main_dataset$surveyor)

```
### b. recoding missing values

```{R}

## for variable 'burglaryyn'
main_dataset$burglaryyn <- str_replace_all(main_dataset$burglaryyn, c("-997" = "Refuse to answer", "-999" = "Don't Know"))

unique(main_dataset$burglaryyn)

## for variable 'vandalismyn'
main_dataset$vandalismyn <- str_replace_all(main_dataset$vandalismyn, c("-999" = "Don't Know"))
unique(main_dataset$vandalismyn)

## for variable 'trespassingyn'
main_dataset$trespassingyn <- str_replace_all(main_dataset$trespassingyn, c("-997" = "Refuse to answer"))
unique(main_dataset$trespassingyn)

```

```{r Part II}

```

# Part 2

## 1.) Crop Insurance Project

```{r 1. crop insurance project}
# Loading data

farmer_ind <- read_dta("~/Desktop/RA_test/coding_challenge/Part 2/A/farmer_ind.dta")
income <- read_dta("~/Desktop/RA_test/coding_challenge/Part 2/A/income.dta")
mkt_prices <- read_dta("~/Desktop/RA_test/coding_challenge/Part 2/A/mkt_prices.dta")


## merging data

farmer_full <- farmer_ind %>% 
  left_join(income, by = "farmer_id") %>% 
  left_join(mkt_prices, by = "crop_id")
```

### a). Calculating income of cocoa farmers and others

```{R}

income_cocoa_farmers <- mean(farmer_full$income[farmer_full$crop_id == 8])

income_cocoa_farmers

income_other_farmers <- mean(farmer_full$income[farmer_full$crop_id != 8])

income_other_farmers

```

Both the values are numerically same.

```{R}


## checking for the statistical difference
diff_mean <- t.test(farmer_full$income[farmer_full$crop_id == 8], farmer_full$income[farmer_full$crop_id != 8])

diff_mean
```
A p-value of 1 confirms that the difference is statistically zero. The income of farmers who grow cocoa have income statistically equal to those who do not. This provides no statistical support for or against the growing of cocoa. Based on the data no suggestions can be provided to pursue or go against the growing of cocoa.

### b). Estimating the market price of squash.

For each farmer, we have the information on their quantity of a crop sold, their total income, and the market price of the all crops sans squash. We estimate the price of the squash as follows.

We estimate $P_8$ by calculating residual income. Residual income is the income of a farmer from the sale of the crops sans squash.

$$\sum\limits_{1}^{8}P_j\,x_{ij}=M_i $$
or, $$P_8x_{i8}+\sum\limits_{1}^{8}P_j\,x_{ij}=M_i $$
or, $$M_i-\sum\limits_{1}^{8}P_j\,x_{ij}=P_8x_{i8} $$
or, $$M_i^{*}=P_8x_{i8} $$
or, $$M_i^{*}=\hat{P_8}x_{i8} + e_i$$

We then have the estimate $\hat{P_8}$ due to linear regression.






```{r}
## replacing '-99' with NA for variable 'sale_amt'


farmer_full$sale_amt <- str_replace_all(farmer_full$sale_amt, c("-99" = "NA"))
farmer_full$sale_amt <- as.numeric(farmer_full$sale_amt)
```
A large number of farmers do not have 'sale amount' data for several crops. Removing the farmers or removing the individual observations of crops is not feasible. We impute the missing value by the average 'sale amount' or corresponding crops.

```{R}
## creating a varible 'avg_sale_amt' as follows.
farmer_full <- farmer_full %>% group_by(crop_id) %>% mutate(avg_sale_amt = mean(sale_amt, na.rm = T))


```

- step 1: estimate residual income
We estimate the farmers residual $M_i^{*}$ as follows.
'farmer_res' is the data set that contains all the information about crop prices and quantity sold sans squash. After value imputation of the missing values, we calculate residual income as shown above.

```{r}
## farmers residual

farmer_res <- farmer_full %>% select(farmer_id, crop_id, sale_amt, mkt_price, income) %>% filter(crop_id != 2) %>% filter(sale_amt == -99)

```
- step 2: We then regress residual income on the sale amount for squash.

Eventually, we have the estimate $\hat{P_8}$

## 2.) Vitamin supplement Project


```{r 2. Vitamin}
# Loading data

vitamins <- read_dta("~/Desktop/RA_test/coding_challenge/Part 2/B/vitamins.dta")
```
### a. Average solve time

Variable 'time' in data is of format 'MM.SS'. We extract the minute and second part, multiply the minute part with 60 and add the second part. We then have the 'time' variable in seconds.
```{R}
vitamins$ch <- as.character(vitamins$time)
vitamins[c('min', 'sec')] <- str_split_fixed(vitamins$ch, '\\.', 2)

vitamins$sec <- as.numeric(vitamins$sec)
vitamins$sec[is.na(vitamins$sec)] <- 0
vitamins$min <- as.numeric(vitamins$min)

vitamins$tot_sec <- vitamins$min * 60 + vitamins$sec

avg_time <- mean(vitamins$tot_sec)
paste0("average solve time is ", avg_time)
```
### b. Statistical difference between the solve times of treatment and control groups.


```{R}

### different avg times
avg_time_treatment <- mean(vitamins$tot_sec[vitamins$treat == 1])
paste0("average solve time for treatment group is ", avg_time_treatment)


avg_time_control <- mean(vitamins$tot_sec[vitamins$treat == 0])
paste0("average solve time for control group is ", avg_time_control)
```

We check the statistical difference of means of different groups. Due to unequal sizes of the groups of treament and control, and uneuqal 'sd', a two sample welch-test is conducted.

```{R}


diff_test_time <- t.test(vitamins$tot_sec[vitamins$treat == 1], vitamins$tot_sec[vitamins$treat == 0])

diff_test_time
```

We fail to reject the null hypothesis in favor of the alternate hypothesis. Hence, we conclude that true difference in means is equal to 0. The treatment group does not fare well statistically in the test compared to the control group, i.e, we find no evidence of impact of vitamin supplement on the analytical ability. Also, a p-value greater than 0.1 shows that at the 95% or 90% significance level the mathematical difference is likely to be a fluke.

## c. Better supplement

It is likely that 'A' and 'a' refer to the same supplement. Similarly for 'D', 'B', and 'C'. Replacing them appropriately and capitalizing the rest.

```{r}

supp_list <- unique(vitamins$supplement)
supp_list



vitamins$supplement <- str_replace_all(
  vitamins$supplement, c("c" = "C", "d" = "D",
                         "b" = "B", "a" = "A", 
                         "n" = "N", "r" = "R", 
                         " A" = "A"))

```

Since control group were not provided any supplement, their column is blank for supplements.

```{R}



vitamins %>% filter(treat == 1) %>% group_by(supplement) %>% summarise(tot_time_grp = mean(tot_sec)) %>% filter(tot_time_grp == min(tot_time_grp))




```

'R' is the better supplement.













